Ecological Limits to Forecasting: Estimating Uncertainty in Critical Transitions with Deep Learning

### 2022-05-13 manuscript/manuscript.Rmd
intro text

---
title: Estimating Uncertainty for Tipping Points with Deep Learning
authors:
  - name: Marcus Lapeyrolerie
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: mlapeyro@berkeley.edu
  - name: Carl Boettiger
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: cboettig@berkeley.edu (corresponding author)
abstract: |
  1. Meaningful representations of uncertainty are an essential component of effective ecological forecasting.
  2. Machine learning
  3. We examine scenarios in which stochasticity and non-linearity can lead to extreme uncertainty under common ecological models.  
  4. We show that ..

@@

Uncertainty important to forecasting [@Deitze2018] and decision-making [@Schindler2015]

Machine learning is increasingly effective at forecasting, but methods often focus on creation or assessment only of point predictions [@]. 
Because uncertainty can play such an intrinsic role in ecological dynamics [@Boettiger2018; @Coulson2004; @Oveskienen], it is essential that machine learning applications to ecological forecasting are able to address probablistic forecasts adequately.
We evaluate the ability of some of the most promising machine learning methods for probabilistic forecasts relative to traditional statistical and mechanistic approaches applied to several classic models in ecology.

>>>>

---
title: "Ecological Limits to Forecasting: Estimating Uncertainty in Critical Transitions with Deep Learning"
authors:
  - name: Marcus Lapeyrolerie
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: mlapeyro@berkeley.edu
  - name: Carl Boettiger
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: cboettig@berkeley.edu (corresponding author)
abstract: |
  1. Meaningful representations of uncertainty are an essential component of effective ecological forecasting.
  2. Machine learning
  3. We examine scenarios in which stochasticity and non-linearity can lead to extreme uncertainty under common ecological models.  
  4. We show that ..

@@

Forecasting plays an important and rapidly growing role in both testing fundamental understanding of ecological processes and informing ecological applications and conservation decision-making [@Deitze2018; @Schindler2015].
Meanwhile, rapid advances in machine learning have rapidly improved the prevalence and accuracy of short term forecasts in many fields [@something]. 
Will these emerging methods improve the capacity for forecasts in ecological systems as well?
Ecological dynamics are notoriously complex, with uncertainty and non-linearity playing critical roles [@Boettiger2018; @Coulson2004; @Oveskienen2010].
These challenges are nowhere more evident then in _critical transitions_, sudden shifts in the states or patterns of ecosystem dynamics that are simultaneously more important and more difficult to predict than more gradual changes.
Here, we examine several of the best-known examples of critical transitions in ecological systems.
We evaluate the ability of some of the most promising machine learning methods for probabilistic forecasts relative to traditional statistical and mechanistic approaches applied to several classic models in ecology.
